---
title: "data_exploration"
author: "Jinghan Cui"
date: "12/2/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
#install.packages("rstanarm")

# Remove Package
#remove.packages("rstanarm")

library(rstanarm)
library(survival)
library(rstan)
library(StanHeaders)
library(ggplot2)
library(haven)
library(survminer)
library(tidyverse)
#remotes::install_github('stan-dev/rstanarm@feature/survival', build_vignettes = FALSE)

```

```{r}
MFPS_PNAS_SURVIVAL_DATA <- read_dta("dataverse_files/MFPS_PNAS_SURVIVAL_DATA.dta")
summary(MFPS_PNAS_SURVIVAL_DATA)

```

## Data exploration

In survival analysis, the outcome or dependent variable is the *time to event* where some event times are not observed (IE they are censored). 

Here we consider the more common scenario of *right-censoring*. This is the case where the terminating event is not observed. Observations are instead censored at time `t` .

Our first analysis will treat **overall survival** as the event of interest, as opposed to progression-free survival. In this cohort, the overall survival is described by two variables: `os_status` & `os_months`. 

We will start by inspecting these data.
```{r}
which(is.na(MFPS_PNAS_SURVIVAL_DATA))
sum(is.na(MFPS_PNAS_SURVIVAL_DATA))
sapply(MFPS_PNAS_SURVIVAL_DATA, function(x) sum(is.na (x)))
```


```{r}
###MALAWI FAMILY PLANNING STUDY
###R-SCRIPT 5: SURVIVAL PLOT FILE

###MAHESH KARRA
###MAY 10, 2022

###STEP 0: SETTING WORKING DIRECTORY
###To use the correct file directory, replace "C:/Users/mvkarra/Documents/Harvard 2015-2016/POP CENTER/MALAWI FP INTERVENTION/Publications/YEAR 1 ITT ON CONTRACEPTIVE USE/SUBMISSIONS/PNAS/HARVARD DATAVERSE SUBMISSION" with the appropriate file path under which your data is stored.
###Make sure that the Results subfolder is created to store the results from this analysis.
###SET WORKING DIRECTORY


fit <- survfit(Surv(month_next_preg,event_next_preg_24) ~ treatment, data = MFPS_PNAS_SURVIVAL_DATA)
ggsurvplot(fit, data = MFPS_PNAS_SURVIVAL_DATA, pval = TRUE)

fit_cox <- coxph(Surv(month_next_preg,event_next_preg_24) ~ treatment, data = MFPS_PNAS_SURVIVAL_DATA)%>% 
  broom::tidy(exp = TRUE)
aft.fit <- survreg(Surv(month_next_preg,event_next_preg_24) ~ treatment, dist="weibull", data = MFPS_PNAS_SURVIVAL_DATA)%>% 
  broom::tidy(exp = TRUE)
###K-M PLOT
ggsurvplot(
  fit, 
  data = MFPS_PNAS_SURVIVAL_DATA, 
  censor=FALSE,
  conf.int = TRUE,
  risk.table = "nrisk_cumevents",
  risk.table.col = "strata",# Risk table color by groups
  risk.table.height = 0.3, # Useful to change when you have multiple groups
  break.time.by = 4,   
  ylim = c(0.85, 1),
  xlab = "Months since index birthdate",
  ylab = "Probability",
  legend.labs = 
    c("Control", "Treated"),
  pval = TRUE
)

```




prior predictive checks
```{r}
CHAINS <- 4
CORES <- 2
ITER <- 2000
SEED <- 1234
#exponential proportional hazard function that has a constant baseline hazard.
#Setting the prior_PD argument equal to TRUE ensures that we do not condition on the outcome data and instead draw samples only from the prior distributions.
prior.stan.const <- stan_surv(
Surv(month_next_preg, event_next_preg_24) ~ treatment,
data = MFPS_PNAS_SURVIVAL_DATA,
basehaz = "exp",
prior_PD = TRUE,
chains = CHAINS,
cores = CORES,
iter = ITER,
seed = SEED)
prior.stan.const

prior_summary(prior.stan.const)
#marginal prior distributions for log HR
library(bayesplot)
mcmc_intervals(prior.stan.const, pars = c("treatment","(Intercept)"))
#marginal prior distribution for the HR for treatment
mcmc_intervals(prior.stan.const, pars = c("treatment"),
transformations = exp) + vline_at(1)

#use a different prior distribution
prior.stan.const <- update(prior.stan.const,
prior_intercept = normal(0, 1),
prior = normal(0, 0.5))
#marginal prior distributions for log HR
mcmc_intervals(prior.stan.const, pars = c("treatment","(Intercept)"))
#marginal prior distribution for the HR for treatment
mcmc_intervals(prior.stan.const, pars = c("treatment"),
transformations = exp) + vline_at(1)

#a HR around 1 would be plausible for a clinical trial

```
# Model estimation
```{r}
fit.stan.const <- update(prior.stan.const, prior_PD = FALSE)
print(fit.stan.const,digits=3)
#MAD_SD: a measure of the standard deviation of the marginal posterior distribution of the treatment effect. it is based on a scaling of the Median Absolute Deviation (MAD) from the posterior median of the treatment effect.
```

#Prior vs posterior checks

```{r}
require("cowplot")
#intercept and log hazard ratio
plot_grid(
bayesplot_grid(mcmc_intervals(prior.stan.const),
mcmc_intervals(fit.stan.const),
titles = c("Prior", "Posterior"),
xlim = c(-8, 1),
grid_args = list(nrow = 2)),
bayesplot_grid(mcmc_hist(prior.stan.const),
mcmc_hist(fit.stan.const),
titles = c("Prior", "Posterior"),
grid_args = list(nrow = 2)),
ncol = 2
)
#fit a cox model
fit.coxph <- coxph(Surv(month_next_preg, event_next_preg_24) ~ treatment,data = MFPS_PNAS_SURVIVAL_DATA, x = TRUE)
fit.coxph
#hazard ratio
add_cox_hr <- vline_at(exp(coef(fit.coxph)), color = "green")
bayesplot_grid(
mcmc_hist(prior.stan.const,
pars = c("treatment"),
transformations = exp,
binwidth = 0.05) + add_cox_hr,
mcmc_hist(fit.stan.const,
pars = c("treatment"),
transformations = exp,
binwidth = 0.05) + add_cox_hr,
titles = c("Prior", "Posterior"),
xlim = c(0, 3),
grid_args = list(nrow = 2)
)


```


# extend to different baseline ditribution
```{r}
# weibull model
fit.stan.weib <- update(fit.stan.const,
basehaz = "weibull")
# gompertz model
fit.stan.gomp <- update(fit.stan.const,
basehaz = "gompertz")
# cubic m-spline model (with df = 5)
fit.stan.ms5 <- update(fit.stan.const, basehaz = "ms", basehaz_ops = list(df = 5))
# cubic m-spline model (with df = 10)
fit.stan.ms10 <- update(fit.stan.const,
basehaz = "ms",
basehaz_ops = list(df = 10))
# cubic b-spline model (with df = 5)
fit.stan.bs5 <- update(fit.stan.const,
basehaz = "bs",
basehaz_ops = list(df = 5))
# cubic s-spline model (with df = 10)
fit.stan.bs10 <- update(fit.stan.const,
basehaz = "bs",
basehaz_ops = list(df = 10))
# piecewise constant model (with df = 5)
fit.stan.pw5 <- update(fit.stan.const,
basehaz = "ms",
basehaz_ops = list(degree = 0, df = 5))
# piecewise constant model (with df = 10)
fit.stan.pw10 <- update(fit.stan.const,
basehaz = "ms",
basehaz_ops = list(degree = 0, df = 10))
fits_stan <- list(
"Constant" = fit.stan.const,
"Weibull" = fit.stan.weib,
"Gompertz"= fit.stan.gomp,
"MS (df = 5)" = fit.stan.ms5,
"MS (df = 10)" = fit.stan.ms10,
"BS (df = 5)" = fit.stan.bs5,
"BS (df = 10)" = fit.stan.bs10,
"PW (df = 5)" = fit.stan.pw5,
"PW (df = 10)" = fit.stan.pw10
)

# Comparison of estimated baseline hazards
require("purrr")
plots <- map(fits_stan, plot)
bayesplot_grid(
plots = plots,
ylim = c(0, 0.012),
titles = names(fits_stan),
grid_args = list(ncol = 3))
```

#posterior distribution
```{r}
nd <- data.frame(treatment = c("Treatment","Control"))
head(nd)
ps <- posterior_survfit(fit.stan.weib,
newdata = nd,
times = 0,
extrapolate = TRUE)
head(ps)

panel_labels <- c('1' = "Treatment", '2' = "Control")
pps <- plot(ps) +
facet_wrap(~ id, labeller = labeller(id = panel_labels))
pps
```

# Assessing fit of the survival function
```{r}
# define helper function to add knot locations
add_knots <- function(x) {
knots <- x$basehaz$knots
if (is.null(knots))
return(NULL)
geom_vline(xintercept = knots, color = "green", alpha = 0.5)
}
# generate the 'ps_check' plots
plots <- map(fits_stan, ~ (ps_check(.) + add_knots(.) + ggplot2::ylim(0.5,1)))
# combine the plots
bayesplot_grid(
plots = plots,
titles = names(fits_stan),
grid_args = list(ncol = 3))

test <- posterior_survfit(fit.stan.weib,draw=1)
```

Comparison using leave-one-out cross validation

```{r}
loos <- map(fits_stan, loo)
loo_compare(loos)

```

# adjusted model
```{r}
fit.stan.weib.aj <- update(fit.stan.weib, Surv(month_next_preg, event_next_preg_24) ~ treatment+age_group_bl+edu_primary_bl+ever_use_bl+religion_r_bl+ethnicity_r_bl)
print(fit.stan.weib.aj, decimal=3)

loo1 <- loo(fit.stan.weib)
loo2 <- loo(fit.stan.weib.aj)
loo_compare(loo1,loo2)

```





```{r}
bayesplot::bayesplot_grid(plot(stan_surv_exponential),
                          plot(stan_surv_weibull),
                          plot(stan_surv_gompertz),
                    
                          ylim = c(0, 0.012))
```



exponential Hierchical
```{r}
CHAINS <- 4
CORES <- 2
ITER <- 2000
SEED <- 1234

mod_randint_HC <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment + (1 | area_bl),data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "exp",chains = CHAINS,cores = CORES,
seed = SEED,iter = ITER)
mod_randint_HC <- update(mod_randint_HC ,
prior_intercept = normal(0, 1),
prior = normal(0, 0.5))
summary(mod_randint_HC)

```
The model contains a baseline covariate for treatment (0 or 1) as well as a area-specific
intercept to allow for correlation in the event times for patients from the same areas.
We’ve called the model object mod_randint to denote the fact that it includes a area-specific
(random) intercept. Let’s examine the parameter estimates from the model:
```{r}
print(mod_randint_HC, digits=2)
```
We see that the estimated log hazard ratio for treatment ( beta^hat(trt) = -0.49) is about the samw as 
the "true" log hazard ratio used in the simple model (beta^hat(trt) = -0.50).  The estimated baseline hazard rate is exp(−5.54) = 0.0039, which is almost equal to the baseline hazard rate used in simple model. 

```{r}
mod_fixed <- update(mod_randint_HC, formula = Surv(month_next_preg,event_next_preg_24) ~treatment)
loo_fixed <- loo(mod_fixed)
loo_randint <- loo(mod_randint_HC)
loo_compare(loo_fixed, loo_randint)
```

weibull Hierchical
```{r}
CHAINS <- 4
CORES <- 2
ITER <- 2000
SEED <- 1234

mod_randint_HC_weibull <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment + (1 | area_bl),data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "weibull",chains = CHAINS,cores = CORES,
seed = SEED,iter = ITER)
mod_randint_HC_weibull <- update(mod_randint_HC_weibull ,
prior_intercept = normal(0, 1),
prior = normal(0, 0.5))
summary(mod_randint_HC_weibull)

```

```{r}
print(mod_randint_HC_weibull, digits=2)
```

```{r}
mod_fixed_wei <- update(mod_randint_HC_weibull,basehaz = "weibull", formula = Surv(month_next_preg,event_next_preg_24) ~treatment)
loo_fixed_wei <- loo(mod_fixed_wei)
loo_randint_HC_weibull <- loo(mod_randint_HC_weibull)
loo_compare(loo_fixed_wei, loo_randint_HC_weibull)
```

gompertz Hierchical
```{r}
CHAINS <- 4
CORES <- 2
ITER <- 2000
SEED <- 1234

mod_randint_HC_gom <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment + (1 | area_bl),data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "gompertz",chains = CHAINS,cores = CORES,
seed = SEED,iter = ITER)
mod_randint_HC_gom <- update(mod_randint_HC_gom ,
prior_intercept = normal(0, 1),
prior = normal(0, 0.5))
summary(mod_randint_HC_gom)

```

```{r}
print(mod_randint_HC_gom, digits=2)
```

```{r}
mod_fixed_gom <- update(mod_randint_HC_gom, formula = Surv(month_next_preg,event_next_preg_24) ~treatment)
loo_fixed_gom <- loo(mod_fixed_gom)
loo_randint_HC_gom <- loo(mod_randint_HC_gom)
loo_compare(loo_fixed, loo_randint,loo_fixed_wei, loo_randint_HC_weibull,loo_fixed_gom, loo_randint_HC_gom)
```



```{r}
aj_HC_weibull <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment+age_group_bl+edu_primary_bl+ever_use_bl+religion_r_bl+ethnicity_r_bl + (1 | area_bl),data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "weibull",chains = CHAINS,cores = CORES,
seed = SEED,iter = ITER)
aj_HC_weibull <- update(aj_HC_weibull ,
prior_intercept = normal(0, 1),
prior = normal(0, 0.5))
summary(aj_HC_weibull)
print(aj_HC_weibull, decimal=3)

```

```{r}
loo_randint_HC_weibull <- loo(mod_randint_HC_weibull)
loo_aj_HC_weibull <- loo(aj_HC_weibull)
loo_compare(loo_randint_HC_weibull,loo_aj_HC_weibull)
```

```{r}
aj_HC_gompertz<- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment+age_group_bl+edu_primary_bl+ever_use_bl+religion_r_bl+ethnicity_r_bl + (1 | area_bl),data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "gompertz",chains = CHAINS,cores = CORES,
seed = SEED,iter = ITER)
aj_HC_weibull <- update(aj_HC_gompertz ,
prior_intercept = normal(0, 1),
prior = normal(0, 0.5))
summary(aj_HC_gompertz)
print(aj_HC_gompertz, decimal=3)

```

```{r}
loo_randint_HC_gompertz <- loo(mod_randint_HC_gom)
loo_aj_HC_gompertz <- loo(aj_HC_gompertz)
loo_compare(loo_randint_HC_weibull,loo_aj_HC_weibull,loo_randint_HC_gompertz,loo_aj_HC_gompertz)
```
 As we can see in the output the adjusted model is worse
 
