---
title: "bayesian sruvival"
author: "Ana & Jinghan"
date: "2022-11-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Data exploration

In survival analysis, the outcome or dependent variable is the *time to event* where some event times are not observed (IE they are censored). 

Here we consider the more common scenario of *right-censoring*. This is the case where the terminating event is not observed. Observations are instead censored at time `t` .

Our first analysis will treat **overall survival** as the event of interest, as opposed to progression-free survival. In this cohort, the overall survival is described by two variables: `os_status` & `os_months`. 

We will start by inspecting these data.
```{r}
which(is.na(MFPS_PNAS_SURVIVAL_DATA))
sum(is.na(MFPS_PNAS_SURVIVAL_DATA))
sapply(MFPS_PNAS_SURVIVAL_DATA, function(x) sum(is.na (x)))
```


```{r}
###MALAWI FAMILY PLANNING STUDY
###R-SCRIPT 5: SURVIVAL PLOT FILE

###MAHESH KARRA
###MAY 10, 2022

###STEP 0: SETTING WORKING DIRECTORY
###To use the correct file directory, replace "C:/Users/mvkarra/Documents/Harvard 2015-2016/POP CENTER/MALAWI FP INTERVENTION/Publications/YEAR 1 ITT ON CONTRACEPTIVE USE/SUBMISSIONS/PNAS/HARVARD DATAVERSE SUBMISSION" with the appropriate file path under which your data is stored.
###Make sure that the Results subfolder is created to store the results from this analysis.
###SET WORKING DIRECTORY
getwd()


install.packages(c("ggplot2","haven","survminer","survival"))
library(ggplot2)
library(haven)
library("survminer")
require("survival")

MFPS_PNAS_SURVIVAL_DATA <- read_dta("MFPS_PNAS_SURVIVAL_DATA.dta")
View(MFPS_PNAS_SURVIVAL_DATA)

fit <- survfit(Surv(month_next_preg,event_next_preg_24) ~ treatment, data = MFPS_PNAS_SURVIVAL_DATA)
ggsurvplot(fit, data = MFPS_PNAS_SURVIVAL_DATA, pval = TRUE)


###K-M PLOT
ggsurvplot(
  fit, 
  data = MFPS_PNAS_SURVIVAL_DATA, 
  censor=FALSE,
  conf.int = TRUE,
  risk.table = "nrisk_cumevents",
  risk.table.col = "strata",# Risk table color by groups
  risk.table.height = 0.3, # Useful to change when you have multiple groups
  break.time.by = 4,   
  ylim = c(0.85, 1),
  xlab = "Months since index birthdate",
  ylab = "Probability",
  legend.labs = 
    c("Control", "Treated"),
  pval = TRUE
)

```

Stan Weibull fit

```{r}
stan_weibull_survival_model_file <- system.file('stan', 'weibull_survival_model.stan', package =  'biostan')
biostan::print_stan_file(stan_weibull_survival_model_file)

```
From the data block, observations where events were observed and censored are handled separately. 
```{r}

stan_weibull_survival_model_code <- biostan::read_stan_file(stan_weibull_survival_model_file)
biostan::print_stan_code(stan_weibull_survival_model_code, section = "data")

```

```{r}
stan_weibull_survival_model_data <-
    list(
        ## Number of event individuals
        Nobs = sum(MFPS_PNAS_SURVIVAL_DATA$event_next_preg_24 == 1),
        ## Number of censored individuals
        Ncen = sum(MFPS_PNAS_SURVIVAL_DATA$event_next_preg_24 == 0),
        ## Number of covariates
        M_bg = 1,
        ## Times for event individuals
        yobs = MFPS_PNAS_SURVIVAL_DATA$month_next_preg[MFPS_PNAS_SURVIVAL_DATA$event_next_preg_24 == 1],
        ## Times for censored individuals
        ycen = MFPS_PNAS_SURVIVAL_DATA$month_next_preg[MFPS_PNAS_SURVIVAL_DATA$event_next_preg_24 == 0],
        ## Covariates for event individuals as a matrix
        Xobs_bg = matrix(as.numeric(MFPS_PNAS_SURVIVAL_DATA$treatment)[MFPS_PNAS_SURVIVAL_DATA$event_next_preg_24 == 1]),
        ## Covariates for censored individuals as a matrix
        Xcen_bg = matrix(as.numeric(MFPS_PNAS_SURVIVAL_DATA$treatment)[MFPS_PNAS_SURVIVAL_DATA$event_next_preg_24  == 0])
        )
stan_weibull_survival_model_data

```

```{r}
stan_weibull_survival_model_fit <-
    rstan::stan(file = stan_weibull_survival_model_file,
                data = stan_weibull_survival_model_data)

```

```{r}
stan_weibull_survival_model_fit

```


=======

The traceplots for the parameters of interest appear to indicate reasonable mixing.

```{r}
rstan::traceplot(stan_weibull_survival_model_fit, par = c("alpha","mu","beta_bg"))

```

```{r}
library(rstanarm)
library(survival)
library(rstan)
library(StanHeaders)

```


```{r}
stan_surv_exponential <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment, data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "exp")
summary(stan_surv_exponential)
prior_summary(stan_surv_exponential)

```

```{r}
stan_surv_weibull <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment, data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "weibull")
summary(stan_surv_weibull)
prior_summary(stan_surv_weibull)
```

```{r}
stan_surv_gompertz <- stan_surv(formula = Surv(month_next_preg,event_next_preg_24) ~treatment, data = MFPS_PNAS_SURVIVAL_DATA,basehaz = "gompertz")
summary(stan_surv_gompertz )
prior_summary(stan_surv_gompertz )
```